Job Demands-Resources model components through the lens of O*NET classifications

### 2022-08-15 SIOP_Onet_FA.Rmd
New .Rmd for JDR interaction

bibliography      : ["r-references.bib"]

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

@@

## Material

>>>>

bibliography      : ["r-references.bib", "SIOP_Onet_FA.bib", "jdr.bib", "jobdemands_resources_bibtex.bib"]

@@

csl               : apa7.csl
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

Given the popularity of the Job Demands-Resources Theory [JD-R; @demerouti2001job] in exploring questions related to everything from motivation to job design, we aim to explore the intersection between *perceptions* of job demands and resources, and the broad set of job characteristics provided on O\*Net. This project makes three contributions. We aim to first explore whether ratings of of O\*Net item groupings align with the stated "resources" and "demands" presented in the job demands-resources theory. We then present evidence documenting whether O\*Net job and task descriptors are similarly rated as resources, challenge- or hindrance demands, and lastly, whether such ratings differ across job categories/classifications. 
Across two studies, a series of evaluations were made that used: 1) direct O\*Net terminology (both descriptor and response option), and 2) JD-R influenced ratings of demand, challenge, or hindrance of different types of workers. Prior to a description of results, a brief overview of both the JD-R theory, the stress appraisal process, and O\*Net is provided.

## The Job demands-Resources Theory

The overarching context for this study is that of the job demands-resources theory, which is an expansion of the well-studied job demands-resources model [@demerouti2001job]. One of the major advantages of the job demands-resources theory is that it allows us to model both work environment and job characteristics via job resources and demands. *Resources* include physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. In contrast, *demands* include components of a job that require sustained effort, and as such, produce psychological or physiological strain (e.g., high work pressure is frequently cited as a common demand; @demerouti2001job). 
Cognitively, the perception of an element of ones job as a resource or demand activates one of two distinct processes: either health impairment (resulting from demands) or motivation [resulting from resources; @bakker2014job]. Pertinent to the current study, demanding job characteristics are frequently associated with negative outcomes [e.g., @bakker2003dual], whereas job characteristics deemed resources have been associated with positive organizational outcomes like engagement and motivation [@bakker2007job].

## Objective vs. Subjective Nature of Demands and Resources: The Role of Appraisal

@searle2015merits note that the majority of the research on workplace demands is based on apriori classifications of demands. However, the stress experience, or process, described early on by @lazarus1984stress is grounded in the assumption that individual appraisals of stressors/demands vary. Their transactional theory of stress and coping states that people continuously appraise stimuli in their environments. An appraisal is the cognitive process whereby meaning is assigned to a stimulus. If a stimulus is appraised as a stressor (threat, challenge, potentially harmful), emotional distress leads to coping of some kind. This action to cope is also associated with another appraisal about the outcome itself and the process continues if the outcomes is not appraised as favorable [@lazarus1984stress]. The stress appraisal process suggests that classifying a job characteristic or environmental condition as an objective demand or resource might be in error.

We next consider the (limited) empirical evidence on this topic. First, some relatively recent research suggests that job demands and resources may not be universally appraised or assigned as such. Starting with job demands, @webster2011extending, for example, studied workload, role ambiguity, and role conflict demands, and found that while each could be appraised primarily as a challenge or hindrance demand, they could also simultaneously be perceived as being both a challenge and hindrance to different degrees. While their study did include resources, it nonetheless points to individual differences on how people perceive stressors at work. Although part of a much larger study on retirement, @sonnega_comparison_2018 compared self-reported (subjective) ratings of degree of physical demand, stress, and need for intense concentration from the Health and Retirement Study with objective ratings from O*Net. Correlations physical demand (r = .52), stress (r = .10), and need for intense concentration (r = .14), again suggesting perhaps that our objective ratings of job demands (and resources) may be subject to a greater level of individual difference than assumed. Next considering resources, @schmitz_interpreting_2019 also captured subjective and objective resources in their study of retirement. Correlations of composite variables for the resources of autonomy (r = .12. p > .01), recognition of work (r = .07, p > .01), and decision freedom (r = .08, p > .01), while significant, certainly do not reflect high levels of overlap.

We do acknowledge as well, that demands and resources are not necessarily consistent across days, or seasons, for many employees. @downes2021incorporating meta-analysis addresses this reality in depth, although it is beyond the scope of this project.

## O\*Net Resource

Originally, the Advisory Panel for the Dictionary of Occupational Titles recommended a system that would “…promote the effective education, training, counseling, and employment of the American workforce. It should accomplish its purpose by providing a database system that identifies, defines, classifies, and describes occupations in the economy in an accessible and flexible manner” (@advisory1993new, p. 6). The result was the now commonly used O\*NET. The Occupational Information Network (O\*NET; onetonline.org) contains a comprehensive description of occupations [@peterson2001understanding]. This widely accessed database houses hundreds of standardized and occupation-specific descriptors most occupations in the US and these descriptions are continually updated. In fact, there was a call to work with experienced I/O psychologists over the summer to update the content for the [Industrial and Organizational Psychologist listing on O\*Net](https://www.onetonline.org/link/summary/19-3032.00). These data, and the tools provided for free on the website (e.g., Career Exploration Tools, “My Next Move for Veterans”, “My Next Move”, Toolkit for Business) are frequently used by counselors, students, human resources departments, and researchers to assist potential applicants discover the skills and training they need for the job of their choice. It is also useful to employers by providing them with information with which to craft job descriptions and help employees determine what skills are needed for promotion.

Of greatest interest here are statements taken from O\*NET ["activity" and "context" classifications](https://www.O\*NETonline.org/find/descriptor/result/4.A.1.b.3) (e.g., items related to information input, interacting with others, physical work conditions, structural job characteristics). One of the first and basic questions is whether or not the categorical examples of "resources" and "demands" described in the Job Demands-Resources Theory [@demerouti2001job], for example, are generally deemed resources or demands as we objectively define them. The next logical question surrounds how "universal" such ratings are. For instance, it is quite possible, given the theoretical and empirical evidence presented above, that there is wide variability in individual appraisal of work activities and context such that some people may rate a given activity as a resource and others a hindrance. A second study extends the findings from Study 1 to a potentially key moderator - job categories/classifications, examining whether ratings of resources, challenge- and hindrance demands differ by job classification.

@@

```{r getdata}

data <- read.csv("SIOPcleaned.csv")      ## from "datacleaning.R" script

## O*NET categories

data$onet.resource.ii <- rowMeans(data[c(162:166)], na.rm=TRUE)
data$onet.resource.mp <- rowMeans(data[c(167:176)], na.rm=TRUE)
data$onet.resource.wo <- rowMeans(data[c(177:185)], na.rm=TRUE)
data$onet.resource.io <- rowMeans(data[c(186:202)], na.rm=TRUE)
data$onet.resource.ir <- rowMeans(data[c(149:161)], na.rm=TRUE)
data$onet.resource.pc <- rowMeans(data[c(128:144,146:148)], na.rm=TRUE)
data$onet.resource.sc <- rowMeans(data[c(118:127)], na.rm=TRUE)

data$overall.resource <- rowMeans(data[c(118:144,147:202)], na.rm=TRUE)

data$onet.hindrance.ii <- rowMeans(data[c(246:247,249:252)], na.rm=TRUE)
data$onet.hindrance.mp <- rowMeans(data[c(252:261)], na.rm=TRUE)
data$onet.hindrance.wo <- rowMeans(data[c(262:270)], na.rm=TRUE)
data$onet.hindrance.io <- rowMeans(data[c(271:287)], na.rm=TRUE)
data$onet.hindrance.ir <- rowMeans(data[c(233:245)], na.rm=TRUE)
data$onet.hindrance.pc <- rowMeans(data[c(213:232)], na.rm=TRUE)
data$onet.hindrance.sc <- rowMeans(data[c(203:212)], na.rm=TRUE)

data$overall.hindrance <- rowMeans(data[c(203:247,249:287)], na.rm=TRUE)

data$onet.challenge.ii <- rowMeans(data[c(332:336)], na.rm=TRUE)
data$onet.challenge.mp <- rowMeans(data[c(337:346)], na.rm=TRUE)
data$onet.challenge.wo <- rowMeans(data[c(347:355)], na.rm=TRUE)
data$onet.challenge.io <- rowMeans(data[c(356:372)], na.rm=TRUE)
data$onet.challenge.ir <- rowMeans(data[c(319:331)], na.rm=TRUE)
data$onet.challenge.pc <- rowMeans(data[c(298:307,309:318)], na.rm=TRUE)
data$onet.challenge.sc <- rowMeans(data[c(288:297)], na.rm=TRUE)

data$overall.challenge <- rowMeans(data[c(288:307,309:372)], na.rm=TRUE)

data$burnout      <- rowMeans(data[c(373:376)], na.rm=TRUE)
data$stress       <- rowMeans(data[c(377:379)], na.rm=TRUE)
data$engagement   <- rowMeans(data[c(380:399)], na.rm=TRUE)        ## added 10/12 because analyses too wonky with 3 subscales

```

There were `r nrow(data)` respondents.

```{r PROCESS}

#Next, try a moderated regression

# PROCESS for R version 4.1.1
# Written by Andrew F. Hayes
# www.afhayes.com
# www.processmacro.org
# Copyright 2012-2022 by Andrew F. Hayes ALL RIGHTS RESERVED
# PROCESS workshop schedule at http://haskayne.ucalgary.ca/CCRAM
#
# Distribution of this code in any form except through processmacro.org 
# is prohibited without the permission of the copyright holder, as is
# distribution after modification.
#
# THIS SOFTWARE IS PROVIDED AS IS, WITHOUT WARRANTY OF ANY KIND
# EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
# MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
# IN NO EVENT SHALL THE COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM
# DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT
# OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
# SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE
# USE OF THIS SOFTWARE IMPLIES AGREEMENT WITH THESE TERMS
#
# To activate, run this script. It may take a few minutes. 
# Patience is a virtue. This will produce a function called process.
# The other functions it creates cannot be accessed by the user
# but are used by the master process function.

process.bcboot3<-function(databcbt,estmte,xp2,badend,priorlo,priorhi)
{
  databcbt<-as.matrix(sort(databcbt))
  badlo<-0;badhi<-0
  pv<-matrix(as.numeric(databcbt < estmte));
  pv<-sum(pv)/nrow(databcbt);ppv<-pv;
  if (pv > .5){ppv<-(1-pv)}
  y5<-sqrt(-2*log(ppv))
  p0<-(-.322232431088);p1<-(-1);p2<-(-.342242088547);p3<-(-.0204231210245)
  p4<-(-.0000453642210148);q0<-(.0993484626060);q1<-(.588581570495)
  q2<-(.531103462366);q3<-(.103537752850);q4<-(.0038560700634)
  xp<-y5+((((y5*p4+p3)*y5+p2)*y5+p1)*y5+p0)/((((y5*q4+q3)*y5+q2)*y5+q1)*y5+q0)
  if (pv <= .5){xp<-(-xp)}
  cilow<-round(nrow(databcbt)*pnorm(2*xp-xp2))
  cihigh<-trunc(nrow(databcbt)*pnorm(2*xp+xp2))+1
  if (cilow < 1){cilow<-1;booterr<-1;badlo<-1}
  if (cihigh > nrow(databcbt)){cihigh<-nrow(databcbt);booterr<-1;badhi<-1}
  llcit<-databcbt[cilow,1]
  ulcit<-databcbt[cihigh,1]
  if ((badlo==1) & (llcit != priorlo)){priorlo<-llcit;badend<-c(badend,llcit)}
  if ((badhi==1) & (ulcit != priorhi)){priorhi<-ulcit;badend<-c(badend,ulcit)}
  bootse<-sd(databcbt)
  cires<-as.matrix(c(bootse,llcit,ulcit))
  cires<-list(cires,badend,priorlo,priorhi)
  return(cires)
}

process.pboot3<-function(databcbt,lcval,hcval)
{
  databcbt<-as.matrix(sort(databcbt))
  llcit<-databcbt[lcval,1]
  ulcit<-databcbt[hcval,1]
  bootse<-sd(databcbt)
  cires<-as.matrix(c(bootse,llcit,ulcit))
  return(cires)
}

process.llrtest3<-function(lm,y,x,b,basemod,iterate,converge)
{
  lm<-as.matrix(lm)
  btemphld<-b
  llrdat<-matrix(-999,nrow(x),(nrow(lm)-sum(lm)))
  llrdf<-ncol(x)-ncol(llrdat)
  llrcnt<-0
  for (llri in (1:nrow(lm)))
  {
    if (lm[llri,1]==0){llrcnt<-llrcnt+1;llrdat[,llrcnt]<-x[,llri]}
  }
  LL2<-process.modelest(y,llrdat,2,0,xp2,5,iterate,converge)
  b<-btemphld
  pvchi<-(1-pchisq((LL2-basemod),df=llrdf))
  fresult<-cbind((LL2-basemod),llrdf,pvchi)
  return(fresult)
}

process.describ3<-function(descdatf,type=0,quantle=1)
{
  desctmp<-matrix(-999,(8-(4*type)),ncol(descdatf))
  # mean, sd, min, max, 16th, 50th, 84th, dich toggle
  for (jd in c(1:ncol(descdatf)))
  {
    descdat<-descdatf[,jd]
    #get the mean, sd, minimum, and maximum */
    desctmp[1,jd]<-mean(descdat)
    desctmp[2,jd]<-sd(descdat)
    desctmp[3,jd]<-min(descdat)
    desctmp[4,jd]<-max(descdat)
    if (type==0)
    {
      minwarn<-0;maxwarn<-0
      tmp=as.numeric(descdat==desctmp[3,jd])+as.numeric(descdat==desctmp[4,jd])
      desctmp[8,jd]<-as.numeric(sum(tmp)==length(tmp))
      if (desctmp[3,jd]==desctmp[4,jd]){desctmp[8,jd]<-2}
      descdat<-matrix(sort(descdat))
      decval<-c(.16,.5,.84)
      for (kd in c(1:3))
      {
        low<-trunc(decval[kd]*(length(descdat)+1))
        lowdec<-decval[kd]*(length(descdat)+1)-low
        value<-descdat[low,1]+(descdat[(low+1),1]-descdat[low,1])*lowdec
        desctmp[(4+kd),jd]<-value
      }
      mnotev<-(1)
      modvals<-matrix(desctmp[5:7,],ncol=ncol(descdatf))
      if (quantle != 1)
      {
        desctmp[5,jd]<-desctmp[1,jd]-desctmp[2,jd]
        desctmp[6,jd]<-desctmp[1,jd]
        desctmp[7,jd]<-desctmp[1,jd]+desctmp[2,jd]
        modvals<-matrix(desctmp[5:7,],ncol=ncol(descdatf))
        mnotev<-(2)
        if (modvals[1,1] < desctmp[3,1]){modvals[1,1]<-desctmp[3,1];minwarn<-1}
        if (modvals[3,1] > desctmp[4,1]){modvals[3,1]<-desctmp[4,1];maxwarn<-1}
      }
      if (desctmp[8,jd]==1)
      {modvals<-matrix(c(desctmp[3,1],desctmp[4,1]))
      mnotev<-0;minwarn<-0;maxwarn<-0
      }
      descrtrn<-list(desctmp,modvals,minwarn,maxwarn,mnotev)
    }
  }
  if (type==1)
  {descrtrn<-list(desctmp)}
  return(descrtrn)
}

process.ftest3<-function(lm,bcoef,cv=0,chr=0,brsq=0,skip=0,y,x)
{
  lmat2<-as.matrix(lm)
  y<-as.matrix(y)
  x<-as.matrix(x)
  n<-nrow(y)
  if (skip==0)
  {
    lmat2<-as.matrix(diag(as.numeric(lm)))
    lmat3<-matrix(0,nrow(lmat2),1)
    for (flp in c(1:ncol(lmat2)))
    {
      if (sum(lmat2[,flp])==1)
      {lmat3<-cbind(as.matrix(lmat3),as.matrix(lmat2[,flp]))}
    }
    lmat2<-as.matrix(lmat3[,2:ncol(lmat3)])
  }
  fratio<-(t(t(lmat2)%*%bcoef)%*%solve(t(lmat2)%*%cv%*%lmat2)%*%((t(lmat2)%*%bcoef)))/ncol(lmat2)
  pfr<-(1-pf(fratio,ncol(lmat2),(n-nrow(bcoef))))
  fresult<-matrix(c(fratio,ncol(lmat2),(n-nrow(bcoef)),pfr),ncol=4)
  if (chr==1)
  {
    lmat3<-as.matrix(1-rowSums(lmat2))
    xfm<-matrix(0,n,sum(lmat3))
    flpc<-1
    for (flp in (1:nrow(lmat3)))
    {
      if (lmat3[flp,1]==1){xfm[,flpc]=x[,flp];flpc<-flpc+1}
    }
    bfm<-solve(t(xfm)%*%xfm)%*%t(xfm)%*%y
    resid<-y-(xfm%*%bfm)
    sstotal<-t(y-(sum(y)/n))%*%(y-(sum(y)/n))
    ssresid<-t(resid)%*%resid
    rsqch<-as.numeric(brsq)-((sstotal-ssresid)/sstotal)
    fresult<-matrix(c(rsqch,fresult),ncol=5)
  }
  ftestout<-as.matrix(fresult)
  return(ftestout)
}

#type1 for ols, type2 for logistic LLR, type3 for logistic bootstrapping
process.modelest<-function(y,x,type,full,xp2,hc=5,iterate=100,converge=.00001)
{
  if (type==1)
  {
    invxtx<-solve(t(x)%*%x)
    b<-invxtx%*%t(x)%*%y
    modres<-b
    if (full==1)
    {
      n1<-nrow(x)
      dfres<-(n1-(ncol(x)))
      sstotal<-t(y-(sum(y)/n1))%*%(y-(sum(y)/n1))
      resid=y-x%*%b
      ssresid<-sum(t(resid)%*%resid)       
      r2<-(sstotal-ssresid)/sstotal
      adjr2<-(1-((1-r2)*(n1-1)/(dfres)))
      mse<-ssresid/(n1-ncol(x))
      #HC covariance matrix
      varb<-mse*invxtx
      k3<-ncol(x)
      xhc<-0
      if (hc != 5)
      {
        xhc<-x
        hat<-matrix(xhc[,1])
        for (i3 in c(1:nrow(xhc)))
        {
          xhcm<-matrix(xhc[i3,])
          hat[i3,1]<-t(xhcm)%*%invxtx%*%xhcm
        }     
        if ((hc==0) | (hc==1))
        {
          for (i3 in c(1:k3)){xhc[,i3]<-xhc[,i3]*resid}
        }
        if ((hc==3) | (hc==2))
        {
          for (i3 in c(1:k3))
          {xhc[,i3]<-(resid/(1-hat)^(1/(4-hc)))*xhc[,i3]}
        }
        if (hc==4)
        {
          hcmn<-matrix(4,n1,2);hcmn[,2]<-(n1*hat)/k3
          minr<-apply(hcmn,1,FUN=min)
          for (i3 in c(1:k3))
          {
            xhc[,i3]<-(resid/(1-hat)^(minr/2))*xhc[,i3]
          }
        }
        varb<-(invxtx%*%t(xhc)%*%xhc%*%invxtx)
        if (hc==1){varb<-(n1/(n1-ncol(x)))*varb}
      }
      seb<-sqrt(diag(varb))
      trat<-b/seb
      p<-2*pt(-abs(trat),df=dfres)
      tval<-sqrt(dfres* (exp((dfres-(5/6))*((xp2/(dfres-(2/3)+(.11/dfres)))*(xp2/(dfres-(2/3)+(.11/dfres)))))-1))
      modres<-matrix(c(modres,seb, trat,p,(b-tval*seb),(b+tval*seb)),ncol=6)
      modresl<-t(matrix(c("coeff","hclab","t","p","LLCI","ULCI")))
      lmat<-diag(ncol(x));lmat<-lmat[,2:ncol(lmat)]
      fratio<-(t(t(lmat)%*%b)%*%solve(t(lmat)%*%varb%*%lmat)%*%((t(lmat)%*%b)))/(ncol(x)-1)
      pfr<-1-pf(fratio,(ncol(x)-1),dfres)
      modsum=matrix(c(sqrt(r2),r2,mse,fratio,(ncol(x)-1),dfres,pfr))
      modsuml=matrix(c("R","R-sq","MSE","hcflab","df1","df2", "p"))
      modretrn<-list(modres,modresl,modsum,modsuml,b,varb,tval,resid)
      return(modretrn)
    }
    if (full==0){return(modres)} 
  }
  #for logistic Y model
  if ((type==2) | (type==3))
  {
    xlp<-x;ylp<-as.matrix(y)
    pt2<-matrix((sum(ylp)/nrow(ylp)),nrow(ylp),1)
    if ((type==2)|(type==3)) {LL3<-(ylp*log(pt2)+(1-ylp)*log(1-pt2))}    
    LL3<-(-2*sum(LL3))
    bt1<-matrix(0,ncol(xlp),1);LL1<-0
    pt1<-matrix(0.5,nrow(ylp),1);pt1lp<-pt1
    for (jjj in (1:iterate))
    {
      xlptmp<-t(xlp)
      vecprb<-(pt1lp*(1-pt1lp))
      for (kkk in (1:ncol(xlp))){xlptmp[kkk,]<-xlptmp[kkk,]*t(vecprb)}
      b<-bt1+solve(xlptmp%*%xlp)%*%t(xlp)%*%(ylp-pt1lp)
      if ((type==2) | (type==3))
      {
        xlpb<-xlp%*%b
        xlpbt<-as.numeric(xlpb > -709.7)
        xlpb709<-((1-xlpbt)*(-709.7))
        xlpb<-((xlpb*xlpbt)+xlpb709)
        pt1lp<-1/(1+exp(-(xlpb)))
      }
      itprob<-sum((pt1lp < .00000001) | (pt1lp > .9999999))
      if (itprob > 0)
      {
        for (kkk in (1:nrow(pt1lp)))
        {
          if (pt1lp[kkk,1] > .9999999){pt1lp[kkk,1]<-.9999999}     
          if (pt1lp[kkk,1] < .00000001){pt1lp[kkk,1]<-.00000001}     
        }
        itprob<-0
      }
      if (itprob==0)
      {
        if ((type==2)|(type==3)){LL<-(ylp*log(pt1lp)+(1-ylp)*log(1-pt1lp))}
        LL2<-(-2*sum(LL))
      }
      if (abs(LL1-LL2) < converge)
      {
        if (full==1)
        {
          xlptmp<-t(xlp)
          vecprb<-(pt1lp*(1-pt1lp))   
          for (kkk in (1:ncol(xlp))){xlptmp[kkk,]<-xlptmp[kkk,]*t(vecprb)} 
          varb<-solve(xlptmp%*%xlp)
          seb<-matrix(sqrt(diag(varb)))
        }
        break
      }
      bt1<-b;LL1<-LL2
    }
    modres<-b
    if (jjj > iterate)
    {
      itprob<-2
      if (booting==0){iterrmod<-1}   
      if (booting==1){bootiter==1}   
      if (itprobtg==0)
      {   
        itprobtg<-1;errcode[errs,1]<-47;errs<-errs+1
        if ((booting==0) & (full==1))
        {
          pt1lpc<-(pt1lp*(1-pt1lp))
          vt1<-diag(pt1lpc)
          varb<-solve(t(xlp)%*%vt1%*%xlp)
          seb<-matrix(sqrt(diag(varb)))
        }
      }
    }
    if (full==1)
    {
      trat<-b/seb
      dfres<-nrow(xlp)   
      p<-2*(1-pnorm(abs(trat)))
      modres<-cbind(modres,seb,trat,p) 
      modres<-matrix(c(modres,(b-xp2*seb),(b+xp2*seb)),ncol=6)
      pvchi<-(1-pchisq((LL3-LL2),df=(nrow(modres)-1)))
      mcF<-(LL3-LL2)/LL3
      cox<-1-exp(-(LL3-LL2)/nrow(xlp))
      nagel<-cox/(1-exp(-(LL3)/nrow(xlp)))
      modsum<-matrix(c(LL2,(LL3-LL2),(nrow(modres)-1),pvchi, mcF,cox,nagel))
      modsuml<-matrix(c("-2LL","ModelLL", "df", "p", "McFadden", "CoxSnell", "Nagelkrk"))
      modresl<-t(matrix(c("coeff","se","Z","p","LLCI","ULCI")))
      modretrn<-list(modres,modresl,modsum,modsuml,b,varb,xp2)
      return(modretrn)
    }
    if ((full==0) & (type==2)){return(LL2)} 
    if ((full==0) & (type==3)){return(modres)}
  }
}

process.dummy3<-function(dd)
{
  uq<-unique(dd)
  uq<-matrix(unlist(uq[1]))
  dummy<-matrix(0,nrow(dd),nrow(uq))
  uq<-matrix(sort(uq))
  uq2<-matrix(seq(1:nrow(uq)))
  for (i in c(1:nrow(dd)))
  {for (j in c(1:nrow(uq)))
  {if (dd[i,1]==uq[j,1])
  {dummy[i,uq2[j,1]]<-1}  
  }
  }
  return(dummy)
}

process.makdummy<-function(dd,method,custcov=0,custcode=999)
{
  dd<-dd[order(dd[,2]),]
  newrow<-dd[,1]
  dd<-dd[,2]
  dd<-as.data.frame(dd)
  dummy<-process.dummy3(dd)
  uq<-unique(dd);uq<-matrix(unlist(uq[1]))
  criterrd<-0;errcode5<-0;errcode4<-0;errcode6<-0
  nvls<-ncol(dummy)
  nnvls<-colSums(dummy)
  mnvls<-min(nnvls)
  nnvls<-matrix(sort(uq))
  conmat1<-1
  if (mnvls < 2){errcode5<-1;criterrd<-1}
  if (nvls > 9){errcode4<-1;criterrd<-1}
  x<-0;dummat<-0
  if (criterrd==0)
  {
    dumok<-1
    if (method > 0)
    {
      x<-dummy[,2:ncol(dummy)]
      nx<-ncol(x)
      minus1<-matrix(-1,1,ncol(x))
      if (method==4)
      {
        for (k in (1:nrow(dd)))
        {shole<-sum(x[k,])
        if (shole==0)
        {x[k,]<-minus1}}
      }
      if ((method==2) | (method==3) | (method==5))
      {
        for (k in (1:nrow(dd)))
        {shole<-sum(x[k,])
        if (shole > 0)
        {
          for (i in (1:ncol(x)))
          {if (x[k,i]==0){x[k,i]<-1} else{break}
          }
        }
        }
        if (method==3)
        {
          conmat1=c(-8,1,1,1,1,1,1,1,1,0,-7,1,1,1,1,1,1,1,0,0,-6,1,1,1,1,1,1,0,0,0,-5,1,1,1,1,1,0,0,0,0,-4,1,1,1,1,0,0,0,0,0,-3,1,1,1,0,0,0,0,0,0,-2,1,1,0,0,0,0,0,0,0,-1,1)
          conmat1<-t(matrix(conmat1,9,8))
          for (i in (1:8)){conmat1[i,]=conmat1[i,]/(10-i)}
          conmat1<-t(conmat1[(10-nvls):8,(10-nvls):9])
          for (k in (1:nrow(dd)))
          {shole<-sum(x[k,]);x[k,]=conmat1[(shole+1),]}
        }
      }
      if (method==5)
      {
        if (ncol(custcode) != (nvls*(nvls-1)))
        {errcode6<-1}
        if (ncol(custcode)==(nvls*(nvls-1)))
        {
          conmat1<-matrix(0,nvls,(nvls-1));cnt<-1
          for (i in (1:nvls))
          {
            for (k in (1:(nvls-1)))
            {conmat1[i,k]<-custcode[1,cnt];cnt<-cnt+1}
          }
          for (k in (1:nrow(dd))){x[k,]<-conmat1[(sum(x[k,])+1),]}   #check this
        }
      }
      xskip<-1
      dummat<-matrix(0,(nx+1),nx)
      dummat[(2:nrow(dummat)),]<-diag(nx)
      if (method==4){dummat[1,]<-minus1}
      if (method==2)
      {for (i in 2:nrow(dummat))
      {for (j in (1:(i-1))){dummat[i,j]<-1}}
      }
      if (method==3){dummat<-conmat1}
      if ((method==5) & (errcode6==0)){dummat<-conmat1}
      dummat<-cbind(nnvls,dummat)
    } 
    x<-cbind(newrow,x)
    x<-x[order(x[,1]),]
  }
  dummrn<-list(x,dummat,nvls,nnvls,errcode4,errcode5,errcode6)
  return(dummrn)
}

process<-function(data,y="xxxxx",x="xxxxx",m="xxxxx",w="xxxxx",z="xxxxx",cov="xxxxx",
                  model=999,converge=.00001,iterate=100,hc=5,jn=0,effsize=0,stand=0,xmtest=0,normal=0,
                  mdichok=0,contrast=999,modelbt=0,matrices=0,covmy=0,covcoeff=0,boot=5000,mc=0,
                  intprobe=0.1,plot=0,total=0,save=0,mcx=0,mcw=0,mcz=0,moments=0,progress=1,
                  bmatrix=-999,wmatrix=-999,zmatrix=-999,wzmatrix=-999,cmatrix=-999,xcatcode=999,
                  wcatcode=999,zcatcode=999,wmodval=999,zmodval=999,center=0,conf=95,seed=-999,
                  decimals=9.4,maxboot=0,modelres=0,bc=0,outscreen=1,activate=0,describe=0,listmiss=0,
                  linsum=-999)
{

  #all this is initiation of variables and matrices
  ranseed<-0;
  if (seed != -999)
  {seed<-trunc(abs(seed));set.seed(seed)}
  wnames<-w;znames<-z;mcerpt<-0;wiscov<-0;ziscov<-0;itprobtg<-0;
  v2tag<-0;maxwwarn<-0;minwwarn<-0;maxzwarn<-0;minzwarn<-0
  toomany<-0;wdich<-0;zdich<-0;wnotev<-0;znotev<-0;singlr<-0
  nxpval<-1;nwpval<-1;nzpval<-1;errs<-1;notes<-1;criterr<-0
  novar<-0;adjust<-0;ncs<-0;serial<-0;sobelok<-0;hasw<-0;ydich<-0;
  hasz<-0;printw<-0;printz<-0;counterf<-0;wmodcust<-0;zmodcust<-0
  booting<-0;bootiter<-0;iterrmod<-0;model<-trunc(model)
  errcode<-matrix(0,100,1);notecode<-matrix(0,100,1)
  iterate<-abs(trunc(iterate));converge=abs(converge);badend<-0;booterr<-0;
  itprobtg<-0;v2tag<-0;maxwwarn<-0;minwwarn<-0;maxzwarn<-0
  minzwarn<-0;toomany<-0;wdich<-0;zdich<-0;wnotev<-0;znotev<-0
  nws<-0;nzs<-0;nms<-0;nys<-0;nxs<-0;maxresm<-9;bc<-as.numeric(bc==1);progress<-as.numeric(progress==1)
  mcxok<-0;mcwok<-0;mczok<-0;xprod<-0;zprod<-0;wprod<-0;modcok<-0;alttotal=0;bc<-as.numeric(bc==1)
  jn<-as.numeric(jn==1);effsize<-as.numeric(effsize==1);maxboots=abs(trunc(maxboot))
  normal<-as.numeric(normal==1);xmtest<-as.numeric(xmtest==1);modelres<-as.numeric(modelres==1)
  stand<-as.numeric(stand==1);outscreen<-as.numeric(outscreen==1);activate=as.numeric(activate==1)
  describe<-as.numeric(describe==1);listmiss<-as.numeric(listmiss==1);
  if (stand==1) {effsize<-1}
  pstog<-0;sobelok<-0;mdichok=as.numeric(mdichok==1)
  resultm<-matrix(99999,1,maxresm)

  linsum<-t(matrix(linsum));
  nlinsum<-ncol(linsum);
  if (linsum[1,1]==-999){nlinsum<-0}
